% ======================================================================= %
%                           Introduction
% ======================================================================= %
{\color{blue}\chapter*{Introduction}}
\addcontentsline{toc}{chapter}{Introduction}

This thesis work is about the interactions between human and humanoid robot as co-workers in the industrial scenarios. By interactions, we started with a non-physical human-robot interaction based on a industrially inspired \textit{Pick-n-Place} task example and try to move towards physical human-robot interactions with an example of human humanoid dual arm bi-directional object handover.

This thesis contributes in the broad field of human robot interactions (specially with humanoid robot) both at a safer distance and physical interaction, namely human-robot interaction(HRI) and physical human robot interaction(pHRI). During an empirical industrial co-worker setting, in one HRI study (Chapter \ref{distinct motor contagion}), we examine the effect of motor contagions induced in participants during and after by the observation of the same movements performed by a human, or a humanoid robot co-worker. While in  another study (Chapter \ref{more than just co-workers}) we systematically varied the robot behavior, and observed whether and how the performance of a human participant is affected by the presence of the humanoid robot. We also investigated the effect of physical form of humanoid robot co-worker where torso and head were covered, and only the moving arm was visible to the human participants. Finally, we compared these behaviors with a human co-worker, and examined how the observed behavioral affects scale with experience of robots. In final pHRI study (Chapter \ref{handover chapter}) we designed intuitive bi-directional object handover routine between human and humanoid robot using whole-body control and locomotion, we also predict and estimate the handover position in advance along with the relative orientation of object or human hand during handover and examined the interaction forces during handover of an unknown mass object as well as the effect of robot's physical form on the timing of object handover routine.


%\paragraph*{Chapter \ref{sota}}
%talk little about each chapter
%
%\paragraph*{Chapter \ref{distinct motor contagion}}
%
%\paragraph*{Chapter \ref{more than just co-workers}}
%
%\paragraph*{Chapter \ref{handover chapter}}


\clearpage
