{\color{blue}\chapter*{\centering{Résumé de la thèse}}}
\addcontentsline{toc}{chapter}{Résumé de la thèse}
\thispagestyle{empty}

Human-robot interaction is an emerging field which deals with the study and research of interactions between humans and robots~\cite{goodrich2008human}. The work done in this thesis is about the interactions between human and humanoid robot HRP-2Kai as co-workers in the industrial scenarios. In the context of \textit{non-physical} human-robot interactions (HRI), the studies conducted in the 1$^\text{st}$ part of this thesis investigated the implicit behavioral effects of humanoid robot on the behavior of human co-workers during an industrially inspired \textit{Pick-n-Place} task paradigm. In the context of \textit{physical} human-robot interactions (pHRI), we developed a novel bi-manual object handover framework using robot whole-body control and locomotion in the 2$^\text{nd}$ part of this thesis.\\

%%% roman
\paragraph*{\LARGE {Distinct motor contagions \\}\\}

When an individual (human and robot) performs an action followed by the observation of someone's action, behavioral implicit effects such as motor contagions causes certain features (kinematics parameters, goal or outcome) of that action to become similar to the observed action. However previous studies on motor contagions have examined  effects induced either during the observation of action or after but never together therefore it remains unclear weather and how these effects are distinct from each other.

In Chapter~\ref{distinct motor contagion}, during an industrially inspired repetitive \textit{Pick-n-Place} movement task paradigm, we examined the effect of motor contagions induced in participants during (we call it \textit{on-line} contagions) and after (\textit{off-line} contagions) the observation of the same movements performed by a human, or a humanoid robot co-worker. We specifically examined three important questions:

\begin{enumerate}
	\item Can on-line and off-line contagions from the observation of a same movement affect different movement features of the human participant?
	\item How do the strengths of the on-line and off-line contagions vary with the nature of the co-worker (ie. if human or robot), and the behavior of the co-worker?
	\item Consequently, are the on-line and off-line contagions different, or do they constitute the same effect observed at different instances?
\end{enumerate}

Our results and findings suggest that \textit{on-line} contagions affect participant's movement frequency while the \textit{off-line} contagions affect their movement velocity. Also \textit{off-line} motor contagions were mainly notable after observing human co-worker but the effects of \textit{on-line} contagions were equal with both human and humanoid robot co-workers. Therefore, perhaps the \textit{off-line} contagion is more sensitive to the nature of the co-worker. These two contagions were also observed to be sensitive to the behavioral features of both co-workers, but with robot co-worker, these motor contagions were induced only when robot movement's were biological in nature.  Finally, the overall observations made in this Chapter emphasize on our hypothesis that distinct motor contagions are induced in human participant's \emph{during} the observation of a co-worker (\textit{on-line} contagions) and as well as \emph{after} the observations of same co-worker (\textit{off-line} contagions).



%%% plosone
\paragraph*{\LARGE {Motor contagion influences human co-worker performance \\}\\}

Does the presence of a humanoid robot co-worker influence the performance of humans around it? While the past studies have examined how the effects of induced motor contagions due to the observation of human and robot movements have affected either human co-worker's movement velocity or how it affected movement variance but never both together. Therefore we argue that since precision in movements along with speed is the key in most industrial tasks, hence it is necessary to consider both task accuracy and task speed to accurately measure the performance in a task.

In Chapter~\ref{more than just co-workers}, under the same paradigm and repetitive industrial task and along with the addition of few more conditions, we systematically varied the robot behavior, and observed whether and how the performance of a human participant is affected by the presence of a humanoid co-worker. We also investigated the effect of physical form of humanoid robot where torso and head were covered, and only the moving arm was visible to the human participants. Later, we compared these behaviors with a human co-worker, and examined how the observed behavioral affects scale with experience of robots. 

Primarily, our findings suggest that the presence of a humanoid robot co-worker (or a human co-worker) can influence the performance frequencies of human participants. We observed that participants become slower with a slower co-worker, but also faster with faster co-worker. We also argued that the performance has to be measured considering together, both the task speed (or frequency) as well as task accuracy. We showed how the touch accuracy of the participants have changed alongside the contagions in their {\it htp} and hence the performance of the human co-worker during the task.
 
We also investigated the effects of physical form, by adding two conditions where both human and robot co-worker's head and torso were covered, and human participants were only able to see visible moving arm of the co-worker. Our findings suggest that the presence of a humanoid co-worker can affect human performance, but only when it's humanoid form is visible. Moreover, this effect was supposedly increased with the human participants having prior robot experience. Finally, our results show that human task frequency, but not task accuracy, is affected by the observation of a humanoid robot co-worker, provided the robot's head and torso are visible and robot made biological movements.


%%% handover
\paragraph*{\LARGE {Bi-manual and locomotion synchronized bi-directional object handover \\}\\}

In Chapter~\ref{handover chapter}, we designed a framework to handover an object between human and robot co-workers in the context of pHRI. We concentrated our efforts towards developing a simple yet robust and efficient handover framework. We introduced an intuitive bi-directional object handover \textit{routine} between human-humanoid co-workers using whole-body control (WBC) and locomotion. Throughout this Chapter, the problem of bi-directional object handover between human-humanoid co-workers was treated as one-shot continuous fluid motion. Initially we started by designing a general framework and within it developed models to predict human hand position to converge at the handover location, along with estimating the grasp configuration of object and active human hand during handover \textit{cycle}s. We also designed a model to minimize the interaction forces during the handover of an unknown mass object along with minimizing the overall duration of object handover \textit{routine}. We designed these models to answer three important questions related to human robot object handover ---\textbf{when} (\textit{timing}), \textbf{where} (\textit{position in space}) and \textbf{how} (\textit{orientation and interaction forces}) during a handover \textit{routine}.

Within this handover framework, using humanoid robot HRP-2Kai, we first tested and confirmed the feasibility of these models under the scenario where human and robot co-workers always use their right hand and left end-effector respectively. Later a generalized framework was presented and tested where both co-workers were able to choose either of their hand to handover or exchange the object. In addition, thanks to the native low-level controller of our robot HRP-2Kai and by utilizing whole-body configuration, we were able to further extend our handover framework, which allowed robot to use both hands  (bi-manual) simultaneously during the object handover \textit{routine}.

Furthermore for an proactive handover of an object between human and robot, we believed it was important to consider the possibility of robot taking a step to handover or exchange an object with the human co-worker, in scenarios where a short distance travel is required. Hence, we explored the full capabilities of biped humanoid robot and added a scenario where robot needs to proactively take few steps in order to handover or exchange the object between its human co-worker. This scenario had been implemented on HRP-2Kai and tested during both when human-robot dyad uses either single or both hands simultaneously. 


\paragraph*{\LARGE {Conclusion \\}\\}

To conclude this thesis, our achieved results contributed in the broad field of human robot interactions (specially with the humanoid robot) both at a safer distance and in close proximity, namely during a human-robot interaction (HRI) and physical human robot interaction (pHRI) respectively. The work done in this thesis was about the interactions between human and biped humanoid robot as `co-workers' in the industrial scenarios. We started with the non-physical human-robot interaction scenario based on a industrially inspired \textit{Pick-n-Place} task example and then advanced towards the physical human-robot interactions with an example of human humanoid robot dual arm bi-directional object handover.
 
In Chapter~\ref{distinct motor contagion} and~\ref{more than just co-workers}, we examined an empirical repetitive industrial task in which a human participant and a humanoid robot co-worker near each other. We primarily chose cyclic and repetitive \textit{pick-n-place} task for the experiments as we wanted a task that is simple but rich and could represent several industrial co-worker scenarios. We found that this is one of the most common task in current industrial platforms where robots are often employed. Note that this study didn't consider effects of factors such age, physical or behavioral characteristics of the partner co-worker. They may have had indirectly affected these two motor contagions, perhaps its an interesting topic of discussion and need to be explored in future research.

Note that in Chapter~\ref{more than just co-workers}, we quantify the participants level of the robot exposure by a self perceived robot exposure score by themselves rather than the actual robot exposure. As an actual standard robot exposure questionnaire to measure this effect is currently absent and the development of one can be useful to understand how this effects, such as the one we highlight here, vary over time.

Overall the findings mentioned and discussed in Chapters~\ref{distinct motor contagion} and~\ref{more than just co-workers} highlights several new features of motor contagions, but also open new questions for future research. One can find these results useful for customizing the design of robot co-workers in industries and sports in future studies by moderating or exploiting these contagions. While if ethically allowed, these contagions could be utilized to improve co-workers performance speed and hence productivity.


Finally we conclude Chapter~\ref{handover chapter} with a preliminary testing of complete handover framework under above mentioned scenarios including locomotion. We confirm that our bi-directional object handover framework is intuitive and adaptable to several objects of distinguishable physical properties (shape, size and mass), including  industrial tools. It only needs the information of handover object's shape and size, though knowing mass of the object is not important in the beginning. We confirmed the feasibility of our handover framework under several objects with mass ranges from [$0.17$ to $1.1$] kg, during above mentioned scenarios. However, over the course of several handover \textit{routine} trials, we report that the calculated object mass is not accurate and has marginal error of $ \pm10 $\% compared to object's actual mass and needs further improvement on better estimation of the inertial forces at play or  better ways to remove offsets from force sensors. Although it didn't affect the optimal threshold which is important at the time of object handover from robot to human co-worker.

Though overall our method allow us to not stop the motion of end-effector and still being able to handover (both ways) the object, however if the handover occurs while both human and robot end-effectors are moving then this problem of handover would be extend to the problem of object collaboration and manipulation, which is already being studied extensively in our research group. Therefore we concentrated solely on the proactive bi-directional handover problem and hence decided to reduce the end-effectors velocity (${}^{ef}v\simeq0$) at the time of handover. 

By adding a dimension of step-walk locomotion, We did not focus on the problem of motion planning or navigation in a large cluttered environment but instead we concentrated our efforts to solve and optimize object handover problem which requires immediate shared efforts between human-robot dyad in a small space where few steps are necessary and enough for a comfortable and convenient object handover. Our proposed method is simple but effective to take advantage of biped humanoid robot and deal with the problem of bi-directional bi-manual object handover using robot whole-body control and locomotion.



\clearpage % end of resume de la thesis

%In the context of \textit{non-physical} human-robot interactions, the studies conducted in 1$^\text{st}$ part of this thesis (Chapter~\ref{distinct motor contagion} and Chapter~\ref{more than just co-workers}) were related to the behavioral effects of motor contagions and motivated by the `implicit' social interactions between human and humanoid robot co-workers. While in the context of \textit{physical} human-robot interactions, the 2$^\text{nd}$ part of this thesis (Chapter~\ref{handover chapter}) was motivated by the physical manipulations of object and handover between human and humanoid robot co-workers in close proximity using humanoid robot whole-body control framework and locomotion.



%We mainly focused on three important key features during the human humanoid robot object handover routine ---the \textit{timing}(s) of handover, the \textit{pose} of handover and the \textit{magnitude} of the interaction forces between human hand(s) and humanoid robot end-effector(s).


%However, note that in the cases when human co-worker decides to step backward, then for sometime after the forward step-walk is triggered, the interpersonal distance $ D $ would become larger than 1.2 meters (since human is away from the robot), therefore in order for this model to work, the human co-worker shouldn't take more than few steps backward. 
%
%Though at the moment, step-walk is triggered only when interpersonal distance between the co-workers becomes greater than $ 125\% <= D <= 150\% $, however more possibilities including the estimation of direction in which human co-worker may wishes to move (laterally) and handover the object will be considered in the future scope of this study. At last, in near future, a human-user study will be conducted shortly after further improving the performance of the overall proactive handover framework with step-walk locomotion.
